#Analysis of flights delay
## created by Krzysztof Ignasiak
### date :01.04.2019
```{r echo=FALSE}
library(RODBC)
library(formattable)
dbconnection =odbcConnect("flights")
```

**1.Average arrival delay in minutes**
```{r echo=TRUE}
x <- sqlQuery(dbconnection,"SELECT AVG(arr_delay_new) FROM Flight_delays")
format(round(x, 2), nsmall = 2)
```
**2.Maximal arrival delay in hours**
```{r}
sqlQuery(dbconnection,"SELECT ROUND(MAX(arr_delay_new)/60,2) FROM Flight_delays")
```

**3.Information about flights with maximum delay**
```{r}
sqlQuery(dbconnection,"SELECT unique_carrier,origin_city_name,dest_city_name,arr_delay_new 
FROM Flight_delays
Where arr_delay_new = (SELECT MAX(arr_delay_new) FROM Flight_delays)")
```
**4.Average arrival delay for each day of the week**  
in minutes ordered by minimum delay
```{r}
sqlQuery(dbconnection,"SELECT W.weekday_name,ROUND(AVG(F.arr_delay_new),2) AS 'average arrive delay'
        FROM Flight_delays AS F
         INNER JOIN Weekdays AS W
         ON F.day_of_week =W.weekday_id
         GROUP by W.weekday_name
        Order by AVG(F.arr_delay_new) ;")
```
**5 .Average arrival daley of carries, which offers flights to San Francisco**
```{r}
sqlQuery(dbconnection,"SELECT unique_carrier,ROUND(AVG(arr_delay_new),2) AS 'average delay'
         FROM Flight_delays
         WHERE unique_carrier IN 
         (SELECT DISTINCT TOP 30 unique_carrier
           FROM Flight_delays
          WHERE origin_city_name ='San Francisco, CA')
         GROUP BY unique_carrier
         ORDER BY ROUND(AVG(arr_delay_new),2);")
```
**6. Percentage of carriers with regular delays**
```{r}
number_of_regular_delay_carriers <-sqlQuery(dbconnection,"SELECT COUNT(DISTINCT unique_carrier)
FROM Flight_delays
                                            WHERE unique_carrier IN(
                                            SELECT unique_carrier
                                            FROM Flight_delays
                                            GROUP by unique_carrier
                                            HAVING AVG(arr_delay_new)>10)")
number_of_unique_carriers <-sqlQuery(dbconnection,"SELECT COUNT(DISTINCT unique_carrier)
                  FROM Flight_delays;")
regular_percentage_delays <-(number_of_regular_delay_carriers/number_of_unique_carriers)
percent(regular_percentage_delays)

```
**7 Correlation between arrival delay and departure delay**
```{r}
dep_delay <- sqlQuery(dbconnection,"SELECT dep_delay_new FROM Flight_delays")
arr_delay <-  sqlQuery(dbconnection,"SELECT arr_delay_new FROM Flight_delays")
pearson <- cor(dep_delay,arr_delay, use= "complete.obs" , method= "pearson")
print(pearson)
```
**8 Change in average arrival delay for each carrier between 1-23 and 24-31 of July**
```{r}
unique_carrier <-sqlQuery(dbconnection,"SELECT DISTINCT unique_carrier
               FROM Flight_delays")
               
delay_before_23 <-sqlQuery(dbconnection,"SELECT AVG(arr_delay_new)
FROM Flight_delays
               WHERE month = 7
               AND day_of_month <=23
               GROUP By unique_carrier")

delay_after_23 <-sqlQuery(dbconnection,"SELECT AVG(arr_delay_new)
FROM Flight_delays
               WHERE month = 7
               AND day_of_month >=24
               GROUP By unique_carrier")


abs_change<- data.frame()
for (i in 1:nrow(unique_carrier))
{ 
  abs_change[i,1] <- format(round(abs(delay_after_23[i,1]-delay_before_23[i,1]), 2), nsmall = 2) 
}
col_name <-'absChange'

names(abs_change) <-col_name
unique_carrier_abs_change <-data.frame(unique_carrier,abs_change)
print(unique_carrier_abs_change)
```
**9. List of carriers, which flies From SFO to PDX as well as From SFO to EUG **
```{r}
sqlQuery(dbconnection,"SELECT DISTINCT unique_carrier
FROM Flight_delays
WHERE origin ='SFO' AND (dest ='PDX'OR dest ='EUG');
")
```
**10.Information about the flights from Chicago to Stanford after 2PM (place of departure,place of arrive,average delay)**
```{r}
sqlQuery(dbconnection, "SELECT DISTINCT TOP 10 origin,dest, ROUND(AVG(arr_delay_new),2) AS 'Average delay'
FROM Flight_delays
         WHERE 
         ((origin ='MDW' OR Origin ='ORD')
         AND
         (dest ='SFO' OR dest ='SJC' or dest ='OAK'))
         AND
         (crs_dep_time>1400)
         GROUP by origin,dest
         ORDER BY ROUND(AVG(arr_delay_new),2) desc;")
```